Irt plot extra

### 2022-02-16 manuscript.Rmd
Seceond check on Manual's Standard

  An expert panel prepared the initial 48 item pool spanning different light exposure related behaviors. Responses, consisting rating the frequency of engaging in the per-item behavior on a 5-point Likert type scale were  collected in an online survey yielding responses from an international sample (690 completed responses, 74 countries, 28 time zones). Exploratory factor analysis (EFA) on an initial subset of our sample (n=428) rendered a five-factor solution with 25 items (Wearing blue light filters, spending time outdoors, using phone and smart-watch in bed, using light before bedtime, using light in the morning and during daytime). In a confirmatory factor analysis (CFA) performed on an independent subset of participants (n=262), we removed two further items to attain the best fit for the five-factor solution (CFI=0.97, TLI=0.96, RMSEA=0.05, SRMR=0.09). The internal consistency reliability coefficient for the total instrument was, McDonald’s Omega(total)=0.73. Measurement model invariance analysis between native and non-native English speakers showed our model attained the highest level of invariance (residual invariance; CFI=0.95, TLI =0.95, RMSEA=0.05). Lastly, a short form of LEBA (n=18) was developed using Item Response Theory on the complete sample (n=690).

@@

                    A[label = 'Initial Development of 48 Items'] 
                    B[label = 'Large Scale Survey (n=690)']
                    C[label = 'Initial Item Analysis']
                    D[label = 'Exploratory Factor Analysis (48 items; n=428)']

                    node[fontname = Helvetica, shape = rectangle, margin = 0.2, width = 16, height = 6, 
                    fixedsize = true, fontsize = 45]
                    E[label = 
                    <Factor Identification<br/>Scree Plot: 6 Factors<br/>Minimum Average Partial Method: 5 Factors<br/>Hull Method: 5 Factors>
                    ]

@@

                     1[label = '5 Factor Solution Retained 25 Items']
                     2[label = <Reliability Analysis on EFA Sample (n=428)<br/>Ordinal alpha F1 =.94; F2 = .76; F3 = .75; F4 =.72; F5 =.62.>]

                     3[label = <Confirmatory Factor Analysis<br/>on the Five Factor Model (25 Items; n=262)<br/>Acceptable Fit ( CFI = .94; TLI .93; RMSEA = .06, SRMR = .11)>]

                    4[label = <Confirmatory Factor Analysis<br/>on the Five Factor Model (23 Items; n=262)<br/>Acceptable Fit ( CFI = .97; TLI= .96; RMSEA = .04, SRMR = .09)>]
                     5[label = <Reliability Analysis (23 Items; n=262)<br/>Ordinal alpha F1=.93; F2=.83; F3 =.70; F4=.69; F5=.52<br/>Mcdonald  omega(total) for the total scale = .73>]
                    8[label = <Measurement Invariance Analysis (n=262)<br/>Residual measurement invariance achieved>]
                    9[label = <Semantic Analysis (n=262):<br/>Semantic Scale Network analysis<br/>Readbility: Grade level 3.33>]
                    10[label = <LEBA Short Form (19 Items; n=690)<br/>Based on Item Response Theory>]
                    }
                subgraph cluster_2 {
                      graph[fontname = Helvetica,shape = rectangle]
                      style = rounded
                     node[fontname = Helvetica, margin = 0.2,width = 18, height = 2.5, 
                fixedsize = true, fontsize = 45 ]
                    6[label = 
                    <6-Factor solution:<br/>one factor emerged with only two salient loadings>
                    ]
                    7[label = 'Rejected']
                    }

@@

Lastly, we sought "Item Response Theory" (IRT) based analysis on developing a short form of LEBA. We fitted each factor of LEBA using the graded response model [@samejima1997handbook] to the combined EFA and CFA sample (n =`r nrow(sem.data)`). IRT assesses the item quality by estimating item discrimination, item difficulty, item information, and test information [@bakerBasicsItemResponse2017]. Item discrimination indicates the pattern of variation in the categorical responses with the changes in latent trait level ($\theta$). Item information curve (IIC) indicates the amount of information an item carries along the latent trait continuum. Here, we reported  the item discrimination parameter and categorize the items according to the suggestions of @bakerBasicsItemResponse2017 : none = 0; very low =0.01 to 0.34; low = 0.35 to 0.64; moderate = 0.65 to 1.34 ; high = 1.35 to 1.69; very high >1.70. We discarded the items with relatively flat item information curve (information <.2) to develop the short form of LEBA. We also assessed the precision of the short LEBA using Test information curve (TIC). TIC  indicates the amount of information an the full-scale carry along the latent trait continuum. 
Item fit and person fit of the fitted IRT models were also analyzed to gather more evidence on validity and meaningfulness of our Tool [@desjardinsHandbookEducationalMeasurement2018]. Item fit was evaluated using  the RMSEA value obtained from Signed-$\chi^2$ index implementation, RMSEA value $\le$.06 was considered adequate item fit. Person fit was estimated using standardized fit index Zh statistics [@drasgow1985appropriateness]. Zh < -2 was be considered as a misfit [@drasgow1985appropriateness].

@@

48 items were generated by an expert panel and  administered to a large sample (n = `r nrow(EFA.descriptives)` to explore the latent structure. Exploratory Factor Analysis revealed a five factor solution with 25 items. ("Wearing blue light filters", "Spending time outdoors", "Using phone and smart-watch in bed", "Using light before bedtime",  and "Using light in the morning and during daytime"). The internal consistency reliability coefficient ordinal alpha ranged between `r apa(range(alpha.tab),2,F)`. As all the retained factors were meaningful and contributed essentially towards our aim we retained all five factors.

A CFA on a separate sample ((n = `r nrow(CFA.descriptives)` gave a five-factor solution (CFI =`r apa(fit.2[5],2,F)`; TLI = `r apa(fit.2[6],2,F)`); RMSEA = `r apa(fit.2[7],2,F)`[`r apa(fit.2[8],2,F)`-`r apa(fit.2[9],2,F)`, 90% CI]) and SRMR = `r apa(fit.2[10],2,F)`) after discarding two item. The internal consistency McDonald's $\omega_t$ of the five factors were  satisfactory (`r apa(rel.k[2,1],2,F)`, `r apa(rel.k[2,2],2,F)`, `r apa(rel.k[2,3],2,F)`, `r apa(rel.k[2,4],2,F)`, `r apa(rel.k[2,5],2,F)`) Internal consistency reliability of the total scale (23 items) was also satisfactory, McDonald's $\omega_t$= `r apa(rel.k[5,6],2,F)`.  In the same sample, our measurement invariance analysis  revealed that the latent structure attained the residual measurement invariance across subgroups: male and female (CFI: `r apa (residual.fitm[5,1],2,F)`, TLI: `r apa (residual.fitm[6,1],2,F)`, SRMR: `r apa (residual.fitm[6,1],2,F)`).

@@

Lastly, we developed a short-LEBA (n=18) using IRT analysis. We fitted a graded response model model to the combined EFA and CFA sample (n =`r nrow(data)`). We discarded 5 items with relatively flat item information curve [I($\theta$) <.20]. IRT analysis indicated short form of LEBA is a psychometrically sound measure. Item fit indexes and person fit index for all five fitted model were acceptable. Items had diverse slope parameters indicating a good range of discrimination- the ability to differentiate respondents with different levels of the light exposure related behavior. Test information curve also indicated a good coverage of underlying trait continuum with precision.

>>>>

  An expert panel prepared the initial 48 item pool spanning different light exposure related behaviors. Responses, consisting rating the frequency of engaging in the per-item behavior on a 5-point Likert type scale were  collected in an online survey yielding responses from a geographically unconstrained sample (690 completed responses, 74 countries, 28 time zones). Exploratory factor analysis (EFA) on an initial subset of our sample (n=428) rendered a five-factor solution with 25 items (Wearing blue light filters, spending time outdoors, using phone and smart-watch in bed, using light before bedtime, using light in the morning and during daytime). In a confirmatory factor analysis (CFA) performed on an independent subset of participants (n=262), we removed two further items to attain the best fit for the five-factor solution (CFI=0.97, TLI=0.96, RMSEA=0.05, SRMR=0.09). The internal consistency reliability coefficient for the total instrument was, McDonald’s Omega(total)=0.73. Measurement model invariance analysis between native and non-native English speakers showed our model attained the highest level of invariance (residual invariance; CFI=0.95, TLI =0.95, RMSEA=0.05). Lastly, a short form of LEBA (n=18) was developed using Item Response Theory on the complete sample (n=690).

@@

                    A[label = 'Initial development of 48 items'] 
                    B[label = 'Large scale survey (n=690)']
                    C[label = 'Initial item analysis']
                    D[label = 'Exploratory factor analysis (48 items; n=428)']

                    node[fontname = Helvetica, shape = rectangle, margin = 0.2, width = 16, height = 6, 
                    fixedsize = true, fontsize = 45]
                    E[label = 
                    <Factor identification<br/>Scree Plot: 6 Factors<br/>Minimum average partial method: 5 factors<br/>Hull method: 5 factors>
                    ]

@@

                     1[label = '5 factor solution retained 25 items']
                     2[label = <Reliability analysis on EFA sample (n=428)<br/>Ordinal alpha F1 =.94; F2 = .76; F3 = .75; F4 =.72; F5 =.62.>]

                     3[label = <Confirmatory factor analysis<br/>on the five factor model (25 items; n=262)<br/>Acceptable fit ( CFI = .94; TLI .93; RMSEA = .06, SRMR = .11)>]

                    4[label = <Confirmatory factor analysis<br/>on the five factor model (23 Items; n=262)<br/>Acceptable fit ( CFI = .97; TLI= .96; RMSEA = .04, SRMR = .09)>]
                     5[label = <Reliability analysis (23 items; n=262)<br/>Ordinal alpha F1=.93; F2=.83; F3 =.70; F4=.69; F5=.52<br/>Mcdonald  omega(total) for the total scale = .73>]
                    8[label = <Measurement invariance analysis (n=262)<br/>Residual measurement invariance achieved>]
                    9[label = <Semantic analysis (n=262):<br/>Semantic scale network analysis<br/>Readbility: grade level 3.33>]
                    10[label = <LEBA short form (18 Items; n=690)<br/>Based on Item Response Theory>]
                    }
                subgraph cluster_2 {
                      graph[fontname = Helvetica,shape = rectangle]
                      style = rounded
                     node[fontname = Helvetica, margin = 0.2,width = 18, height = 2.5, 
                fixedsize = true, fontsize = 45 ]
                    6[label = 
                    <6 factor solution:<br/>One factor emerged with only two salient loadings>
                    ]
                    7[label = 'Rejected']
                    }

@@

Lastly, we sought Item Response Theory (IRT) based analysis on developing a short form of LEBA. We fitted each factor of LEBA using the graded response model [@samejima1997handbook] to the combined EFA and CFA sample (n =`r nrow(sem.data)`). IRT assesses the item quality by estimating item discrimination, item difficulty, item information, and test information [@bakerBasicsItemResponse2017]. Item discrimination indicates the pattern of variation in the categorical responses with the changes in latent trait level ($\theta$). Item information curve (IIC) indicates the amount of information an item carries along the latent trait continuum. Here, we reported  the item discrimination parameter and categorize the items according to the suggestions of @bakerBasicsItemResponse2017 : none = 0; very low =0.01 to 0.34; low = 0.35 to 0.64; moderate = 0.65 to 1.34 ; high = 1.35 to 1.69; very high >1.70. We discarded the items with relatively flat item information curve (information <.2) to develop the short form of LEBA. We also assessed the precision of the short LEBA using Test information curve (TIC). TIC  indicates the amount of information an the full-scale carry along the latent trait continuum. 
Item fit and person fit of the fitted IRT models were also analyzed to gather more evidence on validity and meaningfulness of our Tool [@desjardinsHandbookEducationalMeasurement2018]. Item fit was evaluated using  the RMSEA value obtained from Signed-$\chi^2$ index implementation, RMSEA value $\le$.06 was considered adequate item fit. Person fit was estimated using standardized fit index Zh statistics [@drasgow1985appropriateness]. Zh < -2 was be considered as a misfit [@drasgow1985appropriateness].

@@

48 items were generated by an expert panel and  administered to a large sample (n=`r nrow(EFA.descriptives)` to explore the latent structure. Exploratory Factor Analysis revealed a five factor solution with 25 items. ("Wearing blue light filters", "Spending time outdoors", "Using phone and smart-watch in bed", "Using light before bedtime",  and "Using light in the morning and during daytime"). The internal consistency reliability coefficient ordinal alpha ranged between `r apa(range(alpha.tab),2,F)`. As all the retained factors were meaningful and contributed essentially towards our aim we retained all five factors.

A CFA on a separate sample ((n=`r nrow(CFA.descriptives)` gave a five-factor solution (CFI =`r apa(fit.2[5],2,F)`; TLI = `r apa(fit.2[6],2,F)`); RMSEA = `r apa(fit.2[7],2,F)`[`r apa(fit.2[8],2,F)`-`r apa(fit.2[9],2,F)`, 90% CI]) and SRMR = `r apa(fit.2[10],2,F)`) after discarding two item. The internal consistency McDonald's $\omega_t$ of the five factors were  satisfactory (`r apa(rel.k[2,1],2,F)`, `r apa(rel.k[2,2],2,F)`, `r apa(rel.k[2,3],2,F)`, `r apa(rel.k[2,4],2,F)`, `r apa(rel.k[2,5],2,F)`) Internal consistency reliability of the total scale (23 items) was also satisfactory, McDonald's $\omega_t$= `r apa(rel.k[5,6],2,F)`.In the same sample, our measurement invariance analysis  revealed that the latent structure attained the residual measurement invariance across subgroups: male and female (CFI: `r apa (residual.fitm[5,1],2,F)`, TLI: `r apa (residual.fitm[6,1],2,F)`, SRMR: `r apa (residual.fitm[6,1],2,F)`).

@@

Lastly, we developed a short-LEBA (18 items) using IRT analysis. We fitted a graded response model model to the combined EFA and CFA sample (n=`r nrow(data)`). We discarded five items with relatively flat item information curve [I($\theta$) <.20]. IRT analysis indicated short form of LEBA is a psychometrically sound measure. Item fit indexes and person fit index for all five fitted model were acceptable. Items had diverse slope parameters indicating a good range of discrimination- the ability to differentiate respondents with different levels of the light exposure related behaviour. Test information curve also indicated a good coverage of underlying trait continuum with precision.
